Aspects of RDF Data Constraints in the Social, Behavioural and Economic Sciences

### 2015-04-30 paper/paper7/paper7.tex
evaluation

%\section{Running Example}
%\label{running-example}
\section{Common Vocabularies in SBE Sciences}
\label{rdf-representation}
\tb{jeweils ein satz zu DCAT und XKOS}
%In this section, we describe how data in rectangular format (\emph{PHDD}) and metadata on person-level data sets (\emph{Disco}), aggregated data sets (\emph{QB}), thesauri (\emph{SKOS}), and statistical classifications (\emph{XKOS}) are represented in RDF\footnote{The complete running example in RDF is available at: \url{https://github.com/boschthomas/rdf-validation/tree/master/data/running-example}} and how therefore reused vocabularies are interrelated.

For our evaluation, we examined four different vocabularies commonly used in or developed for the SBE sciences which are briefly introduced in the following. For three of them, we analysed actual data according to constraint violations. For PHDD, there is not yet enough data available so we limit our examination to the vocabulary itself.\ke{Neuer Absatz von mir...}

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
%Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
%\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
%and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

%\textbf{Searching for (Meta)data.}
%\emph{DCAT} enables to represent aggregated and person-level data sets inside of data collections like portals, repositories, catalogs, and archives
%which serve as typical entry points when searching for data.
%Users search for aggregated and person-level data records (\emph{dcat:CatalogRecord}) inside data catalogs (\emph{dcat:Catalog}). 
%As search differs depending on the users’ information need,
%users may only search for records' metadata (e.g., \emph{dcterms:title}, \emph{dcterms:description}), 
%or may formulate more sophisticated queries on aggregated and person-level data sets (\emph{dcat:Dataset}) or their
%distributions (\emph{dcat:Distribution}) which are part of the records. 
%Often, users search for data sets covering particular topics (\emph{dcat:keyword}, \emph{dcat:theme}), time periods (\emph{dcterms:temporal}),  or  locations (\emph{dcterms:spatial}), 
%or for certain formats in which the data distribution is available (\emph{dcterms:format}).

@@

	%\item First, we assigned each constraint type to exactly one of the disjoint sets of constraint types 
%in order to get an overview how many constraint types in relation to the total amount of constraint types are extractable from vocabularies ($\mathcal{CT}_{B}$), are easily definable ($\mathcal{CT}_{S}$), and are rather difficult to specify ($\mathcal{CT}_{C}$). 
In close collaboration with several \emph{SBE} domain experts, we defined 246 constraint 
of 53 different types on six vocabularies (\emph{Disco}, \emph{QB}, \emph{SKOS}, \emph{PHDD}, \emph{DCAT}, \emph{XKOS}) and classified them according to the complexity level of their type and their severity level. 
\ke{Alle vokabulare in section "vocabularies" eingeführt?}\tb{mache ich}
For 114 of these constraints, we evaluated the data quality of 15,694 data sets (4,26 billion triples) of \emph{SBE} research data on three common vocabularies in \emph{SBE} sciences (\emph{Disco}, \emph{QB}, \emph{SKOS}) obtained from 33 SPARQL endpoints.
We distinct two classes of vocabularies: 
(1) well-established vocabularies (e.g., \emph{QB}, \emph{SKOS}) which are widely adopted and accepted and 
(2) newly developed vocabularies (e.g., \emph{Disco} which will be published in 2015) which are either recently published or are still in the publication process.
%It is likely that such a vocabulary is still subject of constant change, that published data may not be consistent with its current version,
%and that early adopters did not properly understand its formal specification.
Based on this evaluation we formulate hypotheses.
Nevertheless, we have to be careful making general statements for all vocabularies 
as these hypotheses still have to be verified or falsified in terms of future work 
by evaluating the quality of data represented by additional well-established and newly developed vocabularies.
\ke{das ist zu früh finde um schon einzuschränken, erst mal ergebnisse, dann hypothesen, dann kritische auseinandersetzung}

>>>>

%\section{Running Example}
%\label{running-example}
\section{Common Vocabularies in SBE Sciences}
\label{rdf-representation}
\tb{Ich habe XKOS und DCAT noch ganz am Ende eingeführt (jeweils nur 1 Satz)}
%In this section, we describe how data in rectangular format (\emph{PHDD}) and metadata on person-level data sets (\emph{Disco}), aggregated data sets (\emph{QB}), thesauri (\emph{SKOS}), and statistical classifications (\emph{XKOS}) are represented in RDF\footnote{The complete running example in RDF is available at: \url{https://github.com/boschthomas/rdf-validation/tree/master/data/running-example}} and how therefore reused vocabularies are interrelated.

For our evaluation, we examined six different vocabularies commonly used in or developed for the SBE sciences which are briefly introduced in the following. For three of them, we analysed actual data according to constraint violations. For PHDD, there is not yet enough data available so we limit our examination to the vocabulary itself.

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
%Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.
\emph{DCAT} enables to represent data sets inside of data collections like portals, repositories, catalogs, and archives
which serve as typical entry points when searching for data.

%\textbf{Searching for (Meta)data.}
%\emph{DCAT} enables to represent data sets inside of data collections like portals, repositories, catalogs, and archives
%which serve as typical entry points when searching for data.
%Users search for aggregated and person-level data records (\emph{dcat:CatalogRecord}) inside data catalogs (\emph{dcat:Catalog}). 
%As search differs depending on the users’ information need,
%users may only search for records' metadata (e.g., \emph{dcterms:title}, \emph{dcterms:description}), 
%or may formulate more sophisticated queries on aggregated and person-level data sets (\emph{dcat:Dataset}) or their
%distributions (\emph{dcat:Distribution}) which are part of the records. 
%Often, users search for data sets covering particular topics (\emph{dcat:keyword}, \emph{dcat:theme}), time periods (\emph{dcterms:temporal}),  or  locations (\emph{dcterms:spatial}), 
%or for certain formats in which the data distribution is available (\emph{dcterms:format}).

@@

	%\item First, we assigned each constraint type to exactly one of the disjoint sets of constraint types 
%in order to get an overview how many constraint types in relation to the total amount of constraint types are extractable from vocabularies ($\mathcal{CT}_{B}$), are easily definable ($\mathcal{CT}_{S}$), and are rather difficult to specify ($\mathcal{CT}_{C}$). 
In close collaboration with several \emph{SBE} domain experts, we defined 246 constraint 
of 53 different types on six vocabularies (\emph{Disco}, \emph{QB}, \emph{SKOS}, \emph{PHDD}, \emph{DCAT}, \emph{XKOS}) and classified them according to the complexity level of their type and their severity level. 
For 114 of these constraints, we evaluated the data quality of 15,694 data sets (4,26 billion triples) of \emph{SBE} research data on three common vocabularies in \emph{SBE} sciences (\emph{Disco}, \emph{QB}, \emph{SKOS}) obtained from 33 SPARQL endpoints.
We distinct two classes of vocabularies: 
(1) well-established vocabularies (e.g., \emph{QB}, \emph{SKOS}) which are widely adopted and accepted and 
(2) newly developed vocabularies (e.g., \emph{Disco} which will be published in 2015) which are either recently published or are still in the publication process.
%It is likely that such a vocabulary is still subject of constant change, that published data may not be consistent with its current version,
%and that early adopters did not properly understand its formal specification.
Based on this evaluation we formulate hypotheses.
Nevertheless, we have to be careful making general statements for all vocabularies 
as these hypotheses still have to be verified or falsified in terms of future work 
by evaluating the quality of data represented by additional well-established and newly developed vocabularies.
\ke{das ist zu früh finde um schon einzuschränken, erst mal ergebnisse, dann hypothesen, dann kritische auseinandersetzung}
